Locally Conservative Discontinuous Petrov-Galerkin Finite Elements for Fluid Problems

### 2013-12-03 docs/Truman/LocalConservation/paper/paper.tex
Small changes to local conservation paper

\documentclass[letterpaper]{article}
%TODO: check for periods after equations
%TODO: bold face proper variables
\usepackage[margin=1.0in]{geometry}
\usepackage{authblk}
\usepackage{amsmath, amssymb, amsthm, mathtools}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{color}
\usepackage{subcaption}
\usepackage{morefloats}
\usepackage{pgf,tikz}
\usetikzlibrary{arrows}

@@

\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\norm}[1]{\left\| #1 \right\|}
\newcommand{\snorm}[1]{\left| #1 \right|}
\newcommand{\LRp}[1]{\left( #1 \right)}
\newcommand{\LRs}[1]{\left[ #1 \right]}
\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
\newcommand{\LRc}[1]{\left\{ #1 \right\}}
\newcommand{\tanbui}[2]{\textcolor{blue}{\sout{#1}} \textcolor{red}{#2}}
\newcommand{\Grad} {\ensuremath{\nabla}}
\newcommand{\Div} {\ensuremath{\nabla\cdot}}
\newcommand{\Nel} {\ensuremath{{N^\text{el}}}}
\newcommand{\jump}[1] {\ensuremath{\LRs{\![#1]\!}}}
\newcommand{\uh}{\widehat{u}}
\newcommand{\fnh}{\widehat{f}_n}
\renewcommand{\L}{L^2\LRp{\Omega}}
\newcommand{\pO}{\partial\Omega}
\newcommand{\Gh}{\Gamma_h}
\newcommand{\Gm}{\Gamma_{-}}
\newcommand{\Gp}{\Gamma_{+}}
\newcommand{\Go}{\Gamma_0}
\newcommand{\Oh}{\Omega_h}
\newcommand{\ptl}{{\partial}}
\newcommand{\bfsig}{\mbox{\boldmath $\sigma$}}
\newcommand{\bfv}{\boldsymbol v}
\newcommand{\bfu}{\boldsymbol u}
\newcommand{\bfU}{\boldsymbol U}
\newcommand{\bfV}{\boldsymbol V}
\newcommand{\bfw}{\boldsymbol w}
\newcommand{\bff}{\boldsymbol f}
\newcommand{\bft}{\boldsymbol t}
\newcommand{\bfpsi}{\boldsymbol\psi}
\newcommand{\bflambda}{\boldsymbol\lambda}
\newcommand{\bfn}{\mbox{\boldmath $n$}}
\newcommand{\bfH}{\mbox{\boldmath $H$}}
\newcommand{\HdivK}{\bfH(\text{div},K)}
\newcommand{\HOneK}{H^{1}(K)}
\newcommand{\HOneOmegah}{H^{-1}(\Omega_h)}
\newcommand{\HdivOmegah}{\bfH(\text{div},\Omega_h)}
\newcommand{\vdeltau}{v_{\delta\bs u_h}}
\newcommand{\taudeltau}{\btau_{\delta\bs u_h}}

@@

\section{Intoduction}
The discontinuous Petrov-Galerkin (DPG) method with optimal test functions has
been under active development for convection-diffusion type systems
\cite{DPG1, DPG2, DPG3, DPG5, DemkowiczHeuer, ChanHeuerThanhDemkowicz2012,
MoroNguyenPeraire11}. In this paper, we develop a theory for the locally conservative
formulation of DPG for convection-diffusion type equations and supplement this
with extensive numerical results.

@@

\subsubsection{Robustness Analysis}

@@

Let us start with the inf-sup condition,
\begin{equation}
\sup_{\hat{t}} \frac{| \sum_K \lambda_K \langle \hat{t},1_K \rangle |}{|| \hat{t} ||_E}
\geq  \beta \LRp{\sum_K \mu(K) \lambda_K^2}^{1/2}\,.
\end{equation}
As in the previous analysis, we select for $\hat{t}$ the trace of Raviart-Thomas
interpolant $\bfpsi_h$ of $\bfpsi = R (\sum_K \lambda_K 1_K)$ where $R$ is the right-inverse
of the divergence operator constructed by Costabel and McIntosh. The only change compared
with the previous analysis, is the evaluation of norm of $\hat{t}_h$. For this, we need
to solve the local problems:
\begin{equation}
((v,\bftau), (\delta v, \delta \bftau)_V = \langle \hat{t}, \delta v \rangle_{\ptl K}
= \int_K \text{div} \bfpsi_h \, \delta v = \int_K \text{div} \bfpsi \, \delta v = \int_K \lambda_K \delta v
= \lambda_K (1_K,\delta v)_K  \quad \forall \delta v \in H^1(K)
\label{eq:local}
\end{equation}
We need then an upper bound of the energy norm of $(v_h,\bftau_h)$:
$$
\LRp{\sum_K || (v,\bftau) ||_V^2}^{1/2}\,.
$$
Substituting $(v,\bftau)$ for $(\delta v,\delta\bftau)$ in~(\ref{eq:local}), we get,
\begin{equation}
|| (v,\bftau) ||_V^2 = \lambda_K (1_K, v_K)
\end{equation}
If we have a robust stability estimate:
\begin{equation}
(1_K, v_K) \leq C \mu(K)^{1/2} || (v,\bftau) ||_K
\end{equation}
(i.e. constant $C$ is independent of $\epsilon$) then
\begin{equation}
||  (v,\bftau) ||_V \leq C \mu(K)^{1/2} | \lambda_K |
\end{equation}
and, eventually, as needed,
\begin{equation}
\sum_K || (v,\bftau) ||_V^2  \leq C^2 \sum_K \mu(K) \lambda_K^2
\end{equation}
which leads to the robust estimate of inf-sup constant $\beta$. For example, it is sufficient if
\begin{equation}
|| v ||_{L^2(K)} \leq || (v,\bftau) ||_V \, .
\label{eq:robust_est1}
\end{equation}
Notice that the stability analysis with the energy norm was, in a sense, easier than
with the quotient norm. Only the divergence of the interpolant $\bfpsi_h$ enters~(\ref{eq:local})
and it coincides with the divergence of $\bfpsi$.

@@

With a locally conservative formulation, we can pass in local problem
\eqref{eq:localSolve} with $\alpha\rightarrow0$. The fact that the test
functions will be determined then up to a constant does not matter, for
$\delta\hat t\in\hat F_h^e$, equation
\eqref{eq:conservativeConfusionSystem}$_1$ is orthogonal to constants.
Mathematically, we are dealing with equivalence classes of functions, but in
order to obtain a single function that we can deal with numerically, we
replace the alpha term with a zero mean scaling condition to obtain the new
test norm,
\begin{align}
\min&\left\{\frac{1}{\epsilon},\frac{1}{\mu(K)}\right\}(\taudeltau,\delta\btau)_K
+(\nabla\cdot\taudeltau-\bbeta\cdot\nabla
v,\nabla\cdot\delta\btau-\bbeta\cdot\nabla v)_K\\
&+(\bbeta\cdot\nabla\vdeltau,\bbeta\cdot\nabla \delta v)_K\nonumber
+\epsilon(\nabla\vdeltau,\nabla \delta v)_K
+\frac{1}{\mu(K)^2}\int_K\vdeltau\int_K\delta v
\label{eq:localSolveMod}
\end{align}
where the $\frac{1}{\mu(K)^2}$ coefficient is needed to make the zero mean term
scale correctly since $\int_K\vdeltau$ and $\int_K\delta v$ both scale like
$\mu(K)$. Numerically this helps with the condition
numbers of the local solve.
It is convenient to be able to take $\alpha\rightarrow0$ as we will see in
some later numerical experiments.

>>>>

\documentclass[letterpaper]{article}
%TODO: check for periods after equations
%TODO: bold face proper variables
%TODO: Make sure titles have consistent case
\usepackage[margin=1.0in]{geometry}
\usepackage{authblk}
\usepackage{amsmath, amssymb, amsthm, mathtools}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{color}
\usepackage{subcaption}
\usepackage{morefloats}
\usepackage{pgf,tikz}
\usetikzlibrary{arrows}

@@

\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\norm}[1]{\left\| #1 \right\|}
\newcommand{\snorm}[1]{\left| #1 \right|}
\newcommand{\LRp}[1]{\left( #1 \right)}
\newcommand{\LRs}[1]{\left[ #1 \right]}
\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
\newcommand{\LRc}[1]{\left\{ #1 \right\}}
\newcommand{\LRl}[1]{\left| #1 \right|}
\newcommand{\tanbui}[2]{\textcolor{blue}{\sout{#1}} \textcolor{red}{#2}}
\newcommand{\Grad} {\ensuremath{\nabla}}
\newcommand{\Div} {\ensuremath{\nabla\cdot}}
\newcommand{\Nel} {\ensuremath{{N^\text{el}}}}
\newcommand{\jump}[1] {\ensuremath{\LRs{\![#1]\!}}}
\newcommand{\uh}{\widehat{u}}
\newcommand{\fnh}{\widehat{f}_n}
\renewcommand{\L}{L^2\LRp{\Omega}}
\newcommand{\pO}{\partial\Omega}
\newcommand{\Gh}{\Gamma_h}
\newcommand{\Gm}{\Gamma_{-}}
\newcommand{\Gp}{\Gamma_{+}}
\newcommand{\Go}{\Gamma_0}
\newcommand{\Oh}{\Omega_h}
\newcommand{\ptl}{{\partial}}
\newcommand{\bfsig}{\mbox{\boldmath $\sigma$}}
\newcommand{\bfv}{\boldsymbol v}
\newcommand{\bfu}{\boldsymbol u}
\newcommand{\bfU}{\boldsymbol U}
\newcommand{\bfV}{\boldsymbol V}
\newcommand{\bfw}{\boldsymbol w}
\newcommand{\bff}{\boldsymbol f}
\newcommand{\bft}{\boldsymbol t}
\newcommand{\bfpsi}{\boldsymbol\psi}
\newcommand{\bflambda}{\boldsymbol\lambda}
\newcommand{\bfn}{\mbox{\boldmath $n$}}
\newcommand{\bfH}{\mbox{\boldmath $H$}}
\newcommand{\HdivK}{\bfH(\text{div},K)}
\newcommand{\HOneK}{H^{1}(K)}
\newcommand{\HOneOmegah}{H^{-1}(\Omega_h)}
\newcommand{\HdivOmegah}{\bfH(\text{div},\Omega_h)}
\newcommand{\vdeltau}{v_{\delta\bs u_h}}
\newcommand{\taudeltau}{\btau_{\delta\bs u_h}}

@@

\section{Introduction}
The discontinuous Petrov-Galerkin (DPG) method with optimal test functions has
been under active development for convection-diffusion type systems
\cite{DPG1, DPG2, DPG3, DPG5, DemkowiczHeuer, ChanHeuerThanhDemkowicz2012,
MoroNguyenPeraire11}. In this paper, we develop a theory for the locally conservative
formulation of DPG for convection-diffusion type equations and supplement this
with extensive numerical results.

@@

\subsubsection{Robustness Analysis}
\label{sec:robustnessAnalysis}

@@

Let us start with the inf-sup condition,
\begin{equation}
\sup_{\hat{t}} \frac{| \sum_K \lambda_K \langle \hat{t},1_K \rangle |}{|| \hat{t} ||_E}
\geq  \beta \LRp{\sum_K \mu(K) \lambda_K^2}^{1/2}\,.
\end{equation}
As in the previous analysis, we select for $\hat{t}$ the trace of Raviart-Thomas
interpolant $\bfpsi_h$ of $\bfpsi = R (\sum_K \lambda_K 1_K)$ where $R$ is the right-inverse
of the divergence operator constructed by Costabel and McIntosh. The only change compared
with the previous analysis, is the evaluation of norm of $\hat{t}_h$. For this, we need
to solve the local problems:
\begin{align}
((v,\bftau), (\delta v, \delta \bftau)_V &= \langle \hat{t}, \delta v \rangle_{\ptl K}
= \int_K \text{div} \bfpsi_h \, \delta v = \int_K \text{div} \bfpsi \, \delta v
\nonumber\\
&= \int_K \lambda_K \delta v
= \lambda_K (1_K,\delta v)_K  \quad \forall \delta v \in
H^1(K)\,\forall\delta\bftau\in\bfH(\text{div},K)
\label{eq:local}
\end{align}
We need then an upper bound of the energy norm of $(v_h,\bftau_h)$:
$$
\LRp{\sum_K || (v,\bftau) ||_V^2}^{1/2}\,.
$$
Substituting $(v,\bftau)$ for $(\delta v,\delta\bftau)$ in~(\ref{eq:local}), we get,
\begin{equation}
|| (v,\bftau) ||_V^2 = \lambda_K (1_K, v_K)
\end{equation}
If we have a robust stability estimate:
\begin{equation}
(1_K, v_K) \leq C \mu(K)^{1/2} || (v,\bftau) ||_K
\label{eq:robustEst1}
\end{equation}
(i.e. constant $C$ is independent of $\epsilon$) then
\begin{equation}
||  (v,\bftau) ||_V \leq C \mu(K)^{1/2} | \lambda_K |
\end{equation}
and, eventually, as needed,
\begin{equation}
\sum_K || (v,\bftau) ||_V^2  \leq C^2 \sum_K \mu(K) \lambda_K^2
\end{equation}
which leads to the robust estimate of inf-sup constant $\beta$. For example, it is sufficient if
\begin{equation}
|| v ||_{L^2(K)} \leq || (v,\bftau) ||_V \, .
\label{eq:robust_est1}
\end{equation}
Notice that the stability analysis with the energy norm was, in a sense, easier than
with the quotient norm. Only the divergence of the interpolant $\bfpsi_h$ enters~(\ref{eq:local})
and it coincides with the divergence of $\bfpsi$.

@@

With a locally conservative formulation, we can pass in local problem
\eqref{eq:localSolve} with $\alpha\rightarrow0$. The fact that the test
functions will be determined then up to a constant does not matter, for
$\delta\hat t\in\hat F_h^e$, equation
\eqref{eq:conservativeConfusionSystem}$_1$ is orthogonal to constants.
Mathematically, we are dealing with equivalence classes of functions, but in
order to obtain a single function that we can deal with numerically, we
replace the alpha term with a zero mean scaling condition to obtain the new
test norm,
\begin{align}
\label{eq:localSolveMod}
\min&\left\{\frac{1}{\epsilon},\frac{1}{\mu(K)}\right\}(\taudeltau,\delta\btau)_K
+(\nabla\cdot\taudeltau-\bbeta\cdot\nabla
v,\nabla\cdot\delta\btau-\bbeta\cdot\nabla v)_K\\
&+(\bbeta\cdot\nabla\vdeltau,\bbeta\cdot\nabla \delta v)_K\nonumber
+\epsilon(\nabla\vdeltau,\nabla \delta v)_K
+\frac{1}{\mu(K)}\int_K\vdeltau\int_K\delta v
\end{align}
where the $\frac{1}{\mu(K)}$ coefficient is an arbitrary scaling condition
that doesn't make a difference mathematically, but can affect the condition
number of the actual solve. In practice, we use $\frac{1}{\mu(K)^2}$
since $\int_K\vdeltau$ and $\int_K\delta v$ both scale like
$\mu(K)$, but $\frac{1}{\mu(K)}$ is more convenient more the analysis in the
next section.
It is convenient to be able to take $\alpha\rightarrow0$ as we will see in
some later numerical experiments.

\subsubsection{Proof of Robust Stability Estimate}
In the robustness analysis in Section~\ref{sec:robustnessAnalysis}, we argued
that if we have a robust stability estimate:
\begin{equation}
(1_K, v_K) \leq C \mu(K)^{1/2} || (v,\bftau) ||_K
\tag{\ref{eq:robustEst1}}
\end{equation}
and
\begin{equation}
| ((v,\bftau),(1_K,{\bf 0}))_V | \leq C \mu(K)^{1/2} \, || (v,\bftau) ||_V\,.
\tag{\ref{eq:robust_est2}}
\end{equation}
We now proceed to show that the robust norms we are using satisfy this
requirement.
Consider the inner product from Equation~(\ref{eq:localSolve}), with
$\alpha=1$. We wish to
verify condition (\ref{eq:robustEst1}) with the norm derived from this inner
product on the right hand side. By Cauchy-Schwarz
\begin{equation}
   \int_K v\cdot
   1\leq\mu(K)^{1/2}\norm{v}_{L^2(K)}\leq\mu(K)^{1/2}\norm{(v,\btau)}_K
   \label{eq:robustProof}
\end{equation}
where $\norm{(v,\btau)}_K$ is the norm derived from the inner product.
Condition~\ref{eq:robust_est2} comes out the same since
$$|(
(v,\btau),(1_K,{\bf0}))| =
\sum_K|(1_K,v_K)|\leq\sum_K\mu(K)^{1/2}\norm{v,\btau}_K
$$
element-wise.

Now we wish the perform the same analysis for the modified inner product in
Equation~(\ref{eq:localSolveMod}). In this case,
condition~(\ref{eq:robustEst1}) follows even more naturally as
\begin{equation}
   \int_K v\cdot1\leq\mu(K)^{1/2}\frac{1}{\mu(K)^{1/2}}\LRl{\int_Kv}
   \leq\norm{(v,\btau)}_K
   \label{eq:robustProof2}
\end{equation}
where $\norm{(v,\btau)}$ now refers to the norm generated by inner product
(\ref{eq:localSolveMod}). Condition~(\ref{eq:robust_est2}) follows by the same
reasoning.
